---
title: "KintsuGI Report"
author: "Joe Shaw"
format: pdf
editor: visual
---

```{r}
#| label: packages
#| include: FALSE

library(tidyverse)
library(janitor)

```

# GI sample results

Here are the results for the number of samples tested with the SeqOne pipeline, one result per sample.

```{r}
#| label: gi-sample-results
#| echo: FALSE
#| message: FALSE

gi_csv_cleaned_orps <- read_csv(paste0(config::get("data_folderpath"),
                                       "02_cleaned/gi_csv_cleaned_orps.csv"))

tbl_gi_orps <- gi_csv_cleaned_orps |> 
  group_by(status) |> 
  count() |> 
  rename("Number of samples" = n,
         "GI status" = status) |> 
  adorn_totals()

knitr::kable(tbl_gi_orps)

```

# GI patient results

Here are the results for the number of patients tested with the SeqOne pipeline, one result per patient.

```{r}
#| label: gi-patient-results
#| echo: FALSE
#| message: FALSE

gi_csv_cleaned_orpp <- read_csv(paste0(config::get("data_folderpath"),
                                       "02_cleaned/gi_csv_cleaned_orpp.csv"))

tbl_gi_orpp <- gi_csv_cleaned_orpp |> 
  group_by(status) |> 
  count() |> 
  rename("Number of patients" = n,
         "GI status" = status) |> 
  adorn_totals()

knitr::kable(tbl_gi_orpp)

```

# Germline results

```{r}
#| label: load-data
#| include: FALSE
#| message: FALSE

gi_glvar_joined <- read_csv(paste0(config::get("data_folderpath"),
                                   "03_joined/",
                                   "gi_glvar_joined.csv"))

stopifnot(anyNA(gi_glvar_joined$glvar_headline_result) == FALSE)

stopifnot(anyDuplicated(gi_glvar_joined$nhsno) == 0)

gi_glvar_joined_mod <- gi_glvar_joined |> 
  mutate(gl_snv_gene = str_extract(string = glvar_hgvs_description,
                                pattern = ".*(BRCA1|BRCA2|BRIP1|PALB2|RAD51D|MSH2|MSH6).*",
                                group = 1),
         gl_cnv_gene = str_extract(string = glvar_description,
                                pattern = ".*(BRCA1|BRCA2|BRIP1|PALB2|RAD51D|MSH2|MSH6).*",
                                group = 1),
         gl_gene = case_when(
           is.na(gl_snv_gene) & !is.na(gl_cnv_gene) ~gl_cnv_gene,
           !is.na(gl_snv_gene) & is.na(gl_cnv_gene) ~gl_snv_gene
         ),
         glvar_classification = case_when(
           # Likely pathogenic reduced penetrance
           glvar_hgvs_description == "NM_000059.3(BRCA2):c.9302T>G p.(Leu3101Arg)" ~"Likely pathogenic",
           TRUE ~glvar_classification
         ),
         glvar_classification = factor(glvar_classification,
                                       levels = c("Pathogenic",
                                                  "Likely pathogenic",
                                                  "Uncertain significance")))

gi_glvar_joined_conclusive <- gi_glvar_joined_mod |> 
  filter(status != "Non-conclusive" &
           glvar_headline_result %in% c("No reportable variant(s) detected",
                                        "Reportable variant(s) detected"))

```

There are the results for patients with conclusive results from SeqOne genomic instability (GI) testing and germline variant testing.

```{r}
#| label: tbl-gi-glvar
#| echo: FALSE

tbl_gi_glvar <- gi_glvar_joined_conclusive |>
  group_by(status, glvar_headline_result) |> 
  count() |> 
  rename("SeqOne GI status" = status,
         "Germline result" = glvar_headline_result,
         "Number of patients" = n) |> 
  adorn_totals()

knitr::kable(tbl_gi_glvar)

```

```{r}
#| label: gi-gl-results-variants
#| echo: FALSE

gi_glvar_plot <- gi_glvar_joined_conclusive |> 
  filter(glvar_headline_result == "Reportable variant(s) detected") |> 
  ggplot(aes(x = gl_gene, y = score)) +
  geom_jitter(shape = 21, 
              size = 2,
              width = 0.1, 
              alpha = 0.8,
              aes(fill = glvar_classification)) +
  theme_bw() +
  scale_fill_manual(values = c("#FF3300",
                               "#FF9900",
                               "#CCCCCC")) +
  labs(x = "Germline gene",
       fill = "Variant classification",
       y = "SeqOne GI score",
       title = "GI results for patients with reportable germline variants")

gi_glvar_plot

```

## Patients with germline BRCA variants and negative GI scores

```{r}
#| label: gi-neg-gl-pos
#| echo: FALSE

glvar_neg_gi <- gi_glvar_joined_mod |> 
  filter(glvar_headline_result == "Reportable variant(s) detected") |> 
  filter(status != "Non-conclusive") |> 
  filter(score < 0.5) |> 
  filter(glvar_classification %in% c("Pathogenic", "Likely pathogenic")) |> 
  filter(gl_gene %in% c("BRCA1", "BRCA2")) |> 
  select(nhsno, surname, score, glvar_hgvs_description) |> 
  mutate(score = round(score, 2))

knitr::kable(glvar_neg_gi)

```


